V4046\,Sgr high-resolution 1.3\,mm continuum ALMA observations: a disc with a thin ring

### 2020-10-02 mnras_template.tex
Add files via upload

% Abstract of the paper
\begin{abstract}
   %%%%context
    V4046 Sgr is a nearby pre-main sequence spectroscopic binary system that has a gas-rich circumbinary disc.
    We report on new $\sim20$\,mas resolution Atacama Large Millimeter Array (ALMA) observations of the radio continuum at 1.3\,mm, and compare with archival SPHERE-IRDIS differential polarised imaging (DPI) of scattered-light at H-band (1.65\,$\micron$) to reach conclusions on the disc structure. We use a parametric a radiative transfer model of the circumbinary disc, computed with \textsc{radmc3d} to interpret the available data, including the spectral-energy-distribution, in terms of a physical model. In addition, we explore decompositions of the imaging data in polar coordinates to better constrain the disc radial structure. 
    We report a two ringed structure for the millimetre-size dust population with a narrow, less than 4.17$\pm$0.94\,au-wide inner ring centred at 13.46$\pm$0.43\,au. 
    This ring is the finest reported so far in a gas rich protoplanetary disc. 
    The ring is surrounded by a $\sim$9\,au-wide gap, flanked by a very sharp outer ring starting at 24\,au. The radial profile of the outer ring exhibits an intriguing break at $\sim$35\,au. 
    %The RT model reproduces accurately the available data.
    %We predict a three ringed structure for the small-dust population with a inner ring at $\sim$6\,au hiding under the coronagraph in the SPHERE observations. 
    %The shadow of the secondary reported previously from the scattered-light data is also reproduced by the RT model. 
    %The DPI image reveals strong forward-scattering at 1.65\,$\micron$, which we reproduced with rather large dust grains of no less than 0.4\,$\micron$, hinting at a deficit of very-small-grains. 
    Using the parametric model scale height as an estimated height, we found that the thin inner ring is stable.
    %Deeper ALMA observations may confirm the lopsided 5\,au ring.

@@

Recent observations of circumstellar discs have revolutionised the way we think about protoplanetary discs, but the focus of resolved imaging with the Atacama Large Millimeter Array (ALMA) or with VLT/SPHERE has been mainly been towards the brighter sources. This report is part of a survey of discs around T-Tauri Stars (DARTSS - ref Avenhaus, Perez et al in pret). Here, we present new 1.3\,mm continuum ALMA observation of V4046\,Sgr, which is a remarkable system where we can observe fundamental processes associated with disc evolution.

@@

\citet{dOrazi} also measured the binary phase in the scattered light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. We can then calculate the binary phase at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained due to the slower cooling of larger grains (cite Casassus?).

@@

The simulated 1.3\,mm continuum image displays some clear similarities with the ALMA observation. The model reproduced the two rings that are present in the observation: the faint inner ring and the outer brighter one. The outer ring has an intensity peak at $\sim$30\,au and a break present at $\sim$35\,au.

As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig. \ref{fig:radprofiles}), we can assume that the model provides a reliable approximation of the disc structure, including, particularly for our interest, the dimensions of the previously unseen 1.3\,mm thin inner ring. Accordingly, taking the parametric model values for the large-dust inner ring we can have that it is located at 13.6\,au, has an estimated width of 2.6\,au, an estimated scale height ranging from 0.25 to 0.32\,au and a dust mass of about 3\,M$_{\earth}$.

@@

The key conclusions of our analysis are as follows.
\begin{itemize}
  \item We report a narrow inner ring for the 1.3\,mm continuum at $r$ = 13.46$\pm$0.43\,au and has an estimated width of 4.17$\pm$0.94\,au. The location of this ring is similar to the inner ring observed in the scattered light image, this reveals that the ring includes a considerable mass of millimetre-sized grains of around 3\,M$_{\earth}$. Using the parametric model scale height value ($h= $ 0.28\,au at 13.46\,au) we have that the ring width is $\sim$15 times its estimated height, making it a stable ring.

>>>>

% Abstract of the paper
\begin{abstract}
   %%%%context
    V4046 Sgr is a nearby pre-main sequence spectroscopic binary system that has a gas-rich circumbinary disc.
    We report on new $\sim$20\,mas resolution Atacama Large Millimeter Array (ALMA) observations of the radio continuum at 1.3\,mm, and compare with archival SPHERE-IRDIS differential polarised imaging (DPI) of scattered-light at H-band (1.65\,$\micron$) to reach conclusions on the disc structure. We use a parametric a radiative transfer model of the circumbinary disc, computed with \textsc{radmc3d} to interpret the available data, including the spectral-energy-distribution, in terms of a physical model. In addition, we explore decompositions of the imaging data in polar coordinates to better constrain the disc radial structure. 
    We report a two ringed structure for the millimetre-size dust population with a narrow, less than 4.17$\pm$0.94\,au-wide inner ring located at 13.46$\pm$0.43\,au. 
    This ring is the finest reported so far in a gas rich protoplanetary disc. 
    The ring is surrounded by a $\sim$9\,au-wide gap, flanked by a very sharp outer ring starting at 24\,au. 
    %The radial profile of the outer ring exhibits an intriguing break at $\sim$35\,au. 
    The RT model reproduces accurately the available data.
    %We predict a three ringed structure for the small-dust population with a inner ring at $\sim$6\,au hiding under the coronagraph in the SPHERE observations. 
    %The shadow of the secondary reported previously from the scattered-light data is also reproduced by the RT model. 
    %The DPI image reveals strong forward-scattering at 1.65\,$\micron$, which we reproduced with rather large dust grains of no less than 0.4\,$\micron$, hinting at a deficit of very-small-grains. 
    Using the parametric model scale height as an estimated height, we found that the thin inner ring is stable.
    %Deeper ALMA observations may confirm the lopsided 5\,au ring.

@@

Recent observations of circumstellar discs have revolutionised the way we think about protoplanetary discs, but the focus of resolved imaging with the Atacama Large Millimeter Array (ALMA) or with VLT/SPHERE has been mainly been towards the brighter sources. This report is part of a survey of discs around T-Tauri Stars (DARTSS - ref Avenhaus, Perez et al in pret). Here, we present new 1.3\,mm continuum ALMA observation of V4046\,Sgr, which is a remarkable system where it is possible to observe fundamental processes associated with disc evolution.

@@

\citet{dOrazi} also measured the binary phase in the scattered light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. We can then calculate the binary phase at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained due to the slower cooling of larger grains (cite Casassus?***).

@@

The simulated 1.3\,mm continuum image displays some clear similarities with the ALMA observation. The model reproduced the two rings that are present in the observation: the faint inner ring and the outer brighter one. %The outer ring has an intensity peak at $\sim$30\,au and a break present at $\sim$35\,au. 
As the radial profiles obtained from the simulated images of the model closely resemble those deduced from the observations (Fig. \ref{fig:radprofiles}), we can assume that the model provides a reliable approximation of the disc structure, including, particularly for our interest, the dimensions of the previously unseen 1.3\,mm thin inner ring. Accordingly, taking the parametric model values for the large-dust inner ring we can have that it is located at 13.6\,au, has an estimated width of 2.6\,au, an estimated scale height ranging from 0.25 to 0.32\,au and a dust mass of about 3\,M$_{\earth}$.

@@

The key conclusions of our analysis are as follows.
\begin{itemize}
  \item We report a narrow inner ring for the 1.3\,mm continuum located at 13.46$\pm$0.43\,au from the stars and has an estimated width of 4.17$\pm$0.94\,au. The location of this ring is similar to the inner ring observed in the scattered light image, this reveals that the ring includes a considerable mass of millimetre-sized grains of around 3\,M$_{\earth}$. Using the parametric model scale height value ($h= $ 0.28\,au at 13.46\,au) we have that the ring width is $\sim$15 times its estimated height, making it a stable ring.
